Arbitrary Precision and Arbitrary $n$-Grams

### 2011-03-01 paper/paper.Rnw
paper changes

		We used python's +gmpy+ package to represent the probabilities as
		arbitrary rational numbers (python's integers are already
		arbitrary-precision), and +mpmath+ to perform the final exponentiation
		to calculate the perplexity (the $n$-th root).  The libraries could only
		compute non-integer powers of integers or floats, and integer powers
		were rounded to the nearest integer, and therefore, inaccurate, so
		(very high precision) floats provided the best results.  This final
		stage of computation of perplexity was the only point where imprecise
		mathematics was used.

@@

\subsection*{Perplexity of test set}
We evaluated our model on the test set by computing
the perplexity of the test set. In addition to the variables
above, we ran this test with smoothing on and with smoothing off.

@@

For each combination of the intependent variables, we
generated 23 sentences from the \emph{War and Peace} corpus
and tested their perplexity. We did the same the Shakespeare corpus
but generated only 7 sentences for each treatment combination.

@@

\subsection*{Pelplexity of test set}
For each corpus, perplexity of test set decreased when the models
were run for longer $n$-grams, and the perplexity
leveled off starting at $n=3$ (figure \ref{fig:test}). Probability computation type had no effect.

Smoothing slightly increased perplexity. The increase
was larger for larger $n$.

\begin{figure}
\begin{center}
<<echo=F,fig=T,height=9>>=
test.plot()
@
\end{center}
\caption{\label{fig:test}Perplexities of the test corpora by $n$-gram length and probability computation method for models trained on the respective training corpora.}
\end{figure}

\subsection*{Perplexity of generated sentences}
Perplexity of generated sentences decreased with when longer $n$-grams
were used in the model (figure \ref{fig:gen}). The perplexity
leveled off starting at $n=3$. Probability computation type did not
affect perplexity.

@@

>>>>

		We used python's \verb+gmpy+ package to represent the probabilities as
		arbitrary rational numbers (python's integers are already
		arbitrary-precision), and \verb+mpmath+ to perform the final
		exponentiation to calculate the perplexity (the $n$-th root).  The
		libraries could only compute non-integer powers of integers or floats,
		and integer powers were rounded to the nearest integer, and therefore,
		inaccurate, so (very high precision) floats provided the best results.
		This final stage of computation of perplexity was the only point where
		imprecise mathematics was used. See table \ref{values}

		\begin{table}[p]
		\begin{tabular}{l r l l | l}
			Corpus & n-gram length & smoothing & math & perplexity\\
			\hline
			Shakespeare &  1 &  on &  log &  499.299251351\\
			Shakespeare &  1 &  on &  ap &  499.299251353146\\
			Shakespeare &  1 &  off &  log &  487.213020203\\
			Shakespeare &  1 &  off &  ap &  487.213020203167\\
			War and peace &  1 &  on &  log &  439.45958473\\
			War and peace &  1 &  on &  ap &  439.459584729599\\
			War and peace &  1 &  off &  log &  431.095976221\\
			War and peace &  1 &  off &  ap &  431.095976221135\\
			Shakespeare &  2 &  on &  log &  61.1651128384\\
			Shakespeare &  2 &  on &  ap &  61.1651128383365\\
			Shakespeare &  2 &  off &  log &  53.723243187\\
			Shakespeare &  2 &  off &  ap &  53.7232431870133\\
			War and peace &  2 &  on &  log &  50.5728265194\\
			War and peace &  2 &  on &  ap &  50.5728265193686\\
			War and peace &  2 &  off &  log &  44.4939985115\\
			War and peace &  2 &  off &  ap &  44.4939985115293\\
			Shakespeare &  3 &  on &  log &  42.3740593974\\
			Shakespeare &  3 &  on &  ap &  42.3740593973802\\
			Shakespeare &  3 &  off &  log &  29.644105241\\
			Shakespeare &  3 &  off &  ap &  29.644105240972\\
			War and peace &  3 &  on &  log &  33.5298099459\\
			War and peace &  3 &  on &  ap &  33.529809945862\\
			War and peace &  3 &  off &  log &  23.6337013897\\
			War and peace &  3 &  off &  ap &  23.6337013896575\\
			Shakespeare &  4 &  on &  log &  41.6859932104\\
			Shakespeare &  4 &  on &  ap &  41.6859932104407\\
			Shakespeare &  4 &  off &  log &  25.5850207553\\
			Shakespeare &  4 &  off &  ap &  25.5850207552466\\
			War and peace &  4 &  on &  log &  33.0110361869\\
			War and peace &  4 &  on &  ap &  33.0110361868792\\
			War and peace &  4 &  off &  log &  20.5857956417\\
			War and peace &  4 &  off &  ap &  20.5857956417055\\
			Shakespeare &  5 &  on &  log &  42.4661223625 \\
			Shakespeare &  5 &  on &  ap &  42.4661223624809\\
			Shakespeare &  5 &  off &  log &  25.0344976529\\
			Shakespeare &  5 &  off &  ap &  25.034497652926\\
			War and peace &  5 &  on &  log &  33.6199337718\\
			War and peace &  5 &  on &  ap &  33.6199337717408\\
			War and peace &  5 &  off &  log &  20.0240585504\\
			War and peace &  5 &  off &  ap &  20.0240585503833\\
			Shakespeare &  6 &  on &  log &  42.7513356663 \\
			Shakespeare &  6 &  on &  ap &  42.7513356662635\\
			Shakespeare &  6 &  off &  log &  24.9268403396 \\
			Shakespeare &  6 &  off &  ap &  24.9268403395792\\
		\end{tabular}
		\caption{Perplexity values for various n-gram parameters}
		\label{values}
		\end{table}

@@

\subsection*{Perplexity of test set}
We evaluated our model on the test set by computing
the perplexity of the test set. In addition to the variables
above, we ran this test with smoothing on and with smoothing off.
We hypothesized that smoothing would reduce perplexity.

@@

For each combination of the intependent variables, we
generated 23 sentences from the \emph{War and Peace} corpus
and tested their perplexity. %ADD SHAKESPEARE

@@

\subsection*{Pelplexity of test set}

\subsection*{Perplexity of generated sentences}
Perplexity of generated sentences decreased with increased
\ngram length

@@

\section*{Conclusion}

	Arbitrary precision does not result in more accurate perplexity computations
	over logarithmic computation, at least to floating point precision on the
	output of the perplexity, and such precision is probably dwarfed by noise in
	the inputs.  Such computations, however, take substantially longer to
	compute, so arbitrary precision computations are probably not worth
	performing.

	As expected, lengthening n-grams to moderate values decreases perplexity of
	the test corpus, but increasing the n-gram length beyond 3 did not yield
	benefits, so it appears that trigrams are reasonable for use computing
	perplexity.
